---
title: "Eindopdracht"
author: "Mohammed Al Hor"
date: "2023-01-19"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(simmer)
library(simmer.plot)
library(dplyr)
library(tidyverse)
```

```{r include=FALSE}
set.seed(42)
# Set up the simulation environment
env <- simmer("parcel depot")
env

mu_1 = 5
sd_1 = 1
mu_2 = 8
sd_2 = 2
lambda=10
rate = 1/lambda # the rate is the 1 divided by the time between each parcel


parcel <- trajectory("parcel's path") %>%
  seize("express worker", 1) %>% # checking in of parcels by express worker
  timeout(function() rnorm(1, mu_1, sd_1)) %>% # mean of 5 with sd of 1
  release("express worker", 1) %>%
  
  seize("admin worker", 1) %>% # preparing the parcel for shipping by admin worker
  timeout(function() rnorm(1, mu_2, sd_2)) %>% # mean of 8 with sd of 2
  release("admin worker", 1)

env %>%
  add_resource("express worker", 1) %>%
  add_resource("admin worker", 2) %>%
  add_generator("parcel", parcel, function() rexp(1, rate))

env %>% run(until=6600) # we run the simulation for 6600 minutes (10 working days)
env %>% get_mon_arrivals()
env %>% get_mon_resources()

plot(get_mon_resources(env))
env_per_user <- get_mon_arrivals(env, per_resource = TRUE)
```


## 2 (a) The mean, and standard deviation of the checking time, admin time, wait time before checking, wait time before admin, and total throughput time of a parcel.
      
Express worker mean and sd of checking time in hours:
```{r}
express_worker <- env_per_user %>% filter(resource == 'express worker')

#calculate the mean and sd:
mean(express_worker$activity_time)/60
sd(express_worker$activity_time)/60
```

Express worker mean and sd of wait time in hours:

```{r}
express_worker_wait <- express_worker$end_time - express_worker$start_time - express_worker$activity_time
mean(express_worker_wait)/60
sd(express_worker_wait)/60
```

Admin worker mean and sd of checking time in hours:

```{r}
admin_worker <- env_per_user %>% filter(resource == 'admin worker')
#calculate the mean and sd:
mean(admin_worker$activity_time)/60
sd(admin_worker$activity_time)/60
```

Admin worker mean and sd of wait time in hours:

```{r}
admin_worker_wait <- admin_worker$end_time - admin_worker$start_time - admin_worker$activity_time
mean(express_worker_wait)/60
sd(express_worker_wait)/60
```

Total throughput of a parcel in hours:

```{r}
total_throughput <- env_per_user %>% group_by(name) %>% 
  summarise(min_start_time = min(start_time), max_end_time = max(end_time)) %>%
  mutate(throughput = max_end_time - min_start_time)

mean(total_throughput$throughput)/60
sd(total_throughput$throughput)/60
```


## 2 (b) A density histogram of the total throughput, checking time, admin time, wait time before checking and wait time before admin in the simulation.
All of these values have been converted to hours
Histogram of total throughput:
```{r}
hist(total_throughput$throughput/60)
```

Histogram of checking time
```{r}
hist(express_worker$activity_time/60)
```

Histogram of admin time
```{r}
hist(admin_worker$activity_time/60)
```

Histogram of wait time before checking
```{r}
hist(express_worker_wait/60)
```

Histogram of wait time before admin
```{r}
hist(admin_worker_wait/60)
```


## (c) The proportion of parcels sent out in time. Please also state the simulated amount of parcels and the runtime.

Proportion of parcels sent out on time:
```{r}
shipping_time <- 2*11*60
shipped_in_time <- total_throughput %>% filter(throughput<= shipping_time)
nrow(shipped_in_time)/nrow(total_throughput)*100
```

Total number or parcels simulated:
```{r}
nrow(total_throughput)
```

Total runtime in hours:
```{r}
sum(total_throughput$throughput)/60
```






